---
title: "Warm-up mini-Report: Mosquito Blood Hosts in \n Salt Lake City, Utah"
author: "FILL IN YOUR NAME"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# ABSTRACT

Fill in abstract at the end after we have finished the methods, results, discussion, conclusions and know what our data "says".

# BACKGROUND

Fill in some text here that provides background info on the WNV system, the blood meal DNA extractions, PCR, sequencing, etc. and the foundation for our question/hypothesis.

For example, we can use the viremia duration (Kumar et al., 2003) bar plot (make sure to reference sources!!!)  to illustrate the potential importance of house finches in WNV transmission and as the logical foundation for our hypothesis that house finches serve as amplifying hosts for WNV... and the prediction that locations with more house finches in our blood host analysis are also the same locations with higher positive tests for WNV in mosquito pools... 

NOTE: Examples of data you can plot for the background info at https://github.com/saarman/BIOL3070/

```{r viremia, include=TRUE, eval=TRUE, warning=FALSE, fig.align='left', fig.width=4.75, fig.height=7}
# Manually transcribe duration (mean, lo, hi) from the last table column
duration <- data.frame(
  Bird = c("Canada Goose","Mallard", 
           "American Kestrel","Northern Bobwhite",
           "Japanese Quail","Ring-necked Pheasant",
           "American Coot","Killdeer",
           "Ring-billed Gull","Mourning Dove",
           "Rock Dove","Monk Parakeet",
           "Budgerigar","Great Horned Owl",
           "Northern Flicker","Blue Jay",
           "Black-billed Magpie","American Crow",
           "Fish Crow","American Robin",
           "European Starling","Red-winged Blackbird",
           "Common Grackle","House Finch","House Sparrow"),
  mean = c(4.0,4.0,4.5,4.0,1.3,3.7,4.0,4.5,5.5,3.7,3.2,2.7,1.7,6.0,4.0,
           4.0,5.0,3.8,5.0,4.5,3.2,3.0,3.3,6.0,4.5),
  lo   = c(3,4,4,3,0,3,4,4,4,3,3,1,0,6,3,
           3,5,3,4,4,3,3,3,5,2),
  hi   = c(5,4,5,5,4,4,4,5,7,4,4,4,4,6,5,
           5,5,5,7,5,4,3,4,7,6)
)

# Choose some colors
cols <- c(rainbow(30)[c(10:29,1:5)])  # rainbow colors

# horizontal barplot
par(mar=c(5,12,2,2))  # wider left margin for names
bp <- barplot(duration$mean, horiz=TRUE, names.arg=duration$Bird,
              las=1, col=cols, xlab="Days of detectable viremia", xlim=c(0,7))

# add error bars
arrows(duration$lo, bp, duration$hi, bp,
       angle=90, code=3, length=0.05, col="black", xpd=TRUE)
```

# STUDY QUESTION and HYPOTHESIS

## Questions  
Fill in here, the question we want to answer... e.g. What bird species is acting as WNV amplifying host in Salt Lake City?

## Hypothesis 
Fill in hypothesis... e.g. House finches are acting as important amplifying hosts of WNV in Salt Lake City.
  
## Prediction
Fill in prediction... e.g. If house finches are acting as important amplifying hosts, we predict that trapping locations where mosquitoes feed on house finches will also have higher rates of confirmed WNV in tested mosquito pools.

# MOTHODS

Fill in here, including overview of procedure and methods used for this project.

## Fill in first analysis
```{r first-analysis, include=TRUE, eval=TRUE, warning=FALSE, fig.align='left', fig.width=4.75, fig.height=7}
# put code for analysis here
```

## Fill in second analysis/plot
```{r second-analysis-or-plot, include=TRUE, eval=TRUE, warning=FALSE, fig.align='left', fig.width=4.75, fig.height=7}
# put code for plotting here
```

# DISCUSSION

## Interpretation - fill in analysis

## Interpretation - fill in analysis/plot

# CONCLUSION

# REFERENCES     

1. Komar N, Langevin S, Hinten S, Nemeth N, Edwards E, Hettler D, Davis B, Bowen R, Bunning M. Experimental infection of North American birds with the New York 1999 strain of West Nile virus. Emerg Infect Dis. 2003 Mar;9(3):311-22. https://doi.org/10.3201/eid0903.020628

2. ChatGPT. OpenAI, version Jan 2025. Used as a reference for functions such as plot() and to correct syntax errors. Accessed `r Sys.Date()`.



